Changes in Patterns of Geometric Sign Groupings as a Model of Social Dynamics Across Aurignacian Phases

### 2025-05-27 endofquarterpaper.qmd
edit

One potential identifier of aggregation sites comes from @fig-div-plot, which uses the tabula package [@Frerebeau_2025] to generate the Shannon-Weaver diversity index of sign types per site, and plot the expected diversity index for a given sample size against the actual diversity index per site [@Kintigh_1984]. The plot identifies multiple sites that diverge from the diversity index predicted by sample size, including those with a higher index than expected, such as Hohle Fels in the Transitional phase, Spy, La Ferrassie, and Les Cottés in the Proto-Aurignacian, Solutré, Hohle Fels, and Grottes de Fonds-de-Forêt in the Early Aurignacian, and Vogelherd, Bockstein-Törle, and Vindija Cave in the Evolved Aurignacian. It is possible that those sites with a greater than expected diversity index acted as aggregation sites or served a similar purpose. @fig-neighbor-net also shows La Ferrassie in the Proto-Aurignacian as having multiple connections to sites in both groups, perhaps suggesting that it served as an inter-group aggregation site.

@@

@fig-div-plot shows a great deal of general divergence from the expected diversity index for the sample size. Another potential explanation for the observed variation in diversity is that it may indicate variation in social structure. Despite the tendency to conceive of hunter-gatherers as homogenous, previous studies have established a potential for variation in hunter-gatherer social structures [@Finlayson_Warren_2017; @Riches_1995; @Price_Brown_1985; @Lane_2017; @Pate_2006]. Additional research has established how variation in ornamentation and other forms of material culture can be used to communicate information about social roles and status within groups [@Morwood_1987; @Sauvet_2018; @Mattson_2021; @Larsson_2006; @Moore_2010; @Alfonso-Durrruty_Giles_Misarti2015; @Vanhaeren_2005; @Conkey_1985; @Conkey_1978; @Tilley_1982]. It is therefore possible that greater diversity in geometric signs indicates a wider variety of formal social roles [@Conkey_1985]. This is especially relevant in regards to our larger grouping structure of broader versus restricted signs, as it may indicate two broad types of social structure, with the broader groups having a more formal social structure, and the restricted groups having a less formal social structure.

>>>>

A potential identifier of aggregation sites comes from @fig-div-plot, which uses the tabula package [@Frerebeau_2025] to generate the Shannon-Weaver diversity index of sign types per site, and plot the expected diversity index for a given sample size against the actual diversity index per site [@Kintigh_1984]. The plot identifies multiple sites that diverge from the diversity index predicted by sample size, including those with a higher index than expected, such as Hohle Fels in the Transitional phase, Spy, La Ferrassie, and Les Cottés in the Proto-Aurignacian, Solutré, Hohle Fels, and Grottes de Fonds-de-Forêt in the Early Aurignacian, and Vogelherd, Bockstein-Törle, and Vindija Cave in the Evolved Aurignacian. It is possible that those sites with a greater than expected diversity index acted as aggregation sites or served a similar purpose. La Ferrassie in the Proto-Aurignacian is also shown by @fig-neighbor-net to have multiple connections to sites in both groups, in addition to the diversity index shown in @fig-div-plot, perhaps suggesting that it served as a specifically inter-group aggregation site.

@@

Another potential explanation for the observed general variation in diversity shown in 
@fig-div-plot is that it may indicate variation in social structure. Despite the tendency to conceive of hunter-gatherers as homogeneous, previous studies have established a potential for variation in hunter-gatherer social structures [@Singh_Glowacki_2022; @Finlayson_Warren_2017; @Riches_1995; @Price_Brown_1985; @Lane_2017; @Pate_2006]. Additional research has established how variation in ornamentation and other forms of material culture can be used to communicate information about social roles and status within groups [@Morwood_1987; @Sauvet_2018; @Mattson_2021; @Larsson_2006; @Moore_2010; @Alfonso-Durrruty_Giles_Misarti2015; @Vanhaeren_2005; @Conkey_1985; @Conkey_1978; @Tilley_1982]. It is therefore possible that greater diversity in geometric signs indicates a wider variety of formal social roles [@Conkey_1985]. This is especially relevant in regards to our larger grouping structure of broader versus restricted signs, as it may indicate two broad types of social structure, with the broader groups having a more formal social structure, and the restricted groups having a less formal social structure.
